### 2024-12-11 paper.tex
Initial commit

%%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
%%\documentclass[sigplan,nonacm]{acmart}
\documentclass[sigplan, nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

\graphicspath{{pictures/}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
	\providecommand\BibTeX{{%
			\normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
%%\setcopyright{acmcopyright}
\setcopyright{none}
%%\acmJournal{PACMPL}
%%\acmVolume{1}
%%\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
%%\acmArticle{1}
%%\acmYear{2021}
%%\acmMonth{9}
%%\acmDOI{}
%%\copyrightyear{2018}
%%\acmYear{2018}
%%\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
%%\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium%% on Neural
%%  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
%%\acmPrice{15.00}
%%\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}
\usepackage{xspace}
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.78]{beramono}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{stmaryrd}
%\usepackage{unicode-math}
%\usepackage{MnSymbol}
\usepackage{wasysym}
\usepackage{color}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
%\usepackage[compact]{titlesec}
\usepackage[font={small}]{caption}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{flushend}
\usepackage{bcprules}
\usepackage{textcomp}
\usepackage{tikz}
\usetikzlibrary{positioning,fit,calc,arrows.meta,arrows,decorations}
\usepackage{pdfpages}
\usepackage{cleveref}
\usetikzlibrary{matrix}
\usepackage{xspace}
\definecolor{light-gray}{gray}{0.85}
\usepackage{stackengine}
\usepackage{mathpartir}
\usepackage{float}
\input{macros}

%%
%% end of the preamble, start of the body of the document source.
\newcommand{\rhyme}{\text{Rhyme}\xspace}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\lb}{\{~}
\newcommand{\rb}{~\}}

\newcommand{\Typ}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\Ast}[1]{\ensuremath{\textsf{\text{#1}}}}
\newcommand{\Def}[1]{\ensuremath{\mathrm{#1}}}
\newcommand{\mit}[1]{\ensuremath{\mathit{#1}}}
\newcommand{\msf}[1]{\ensuremath{\mathsf{#1}}}

\newcommand{\lang}{\textsf{SimpLLVM}\xspace}

\newcommand{\Sem}[3][]{\ensuremath{\llbracket {#2} \rrbracket_{#3}^{#1}}}
\newcommand{\SSem}[2]{\ensuremath{\llbracket {#1} \rrbracket_{#2}^{\#}}}
\newcommand{\State}{\mathbb{S}}
\newcommand{\Address}{\mathcal{A}}
\newcommand{\Mem}{\mathbb{M}}
\newcommand{\Value}{\mathcal{V}}
\newcommand{\Loc}{\msf{Loc}}
\begin{document}
\sloppy

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Rhyme: A Formal Query Language for Nested Data Structures}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ruiqi Gao}
\email{gao606@purdue.edu}
\affiliation{%
  \institution{Purdue University}
  \city{Lafayette}
  \state{Indiana}
  \country{USA}
  \postcode{47901}
}
\author{Ran Guo}
\email{guo543@purdue.edu}
\affiliation{%
  \institution{Purdue University}
  \city{Lafayette}
  \state{Indiana}
  \country{USA}
  \postcode{47901}
}
\author{Isaac Fleetwood}
\email{ifleetwo@purdue.edu}
\affiliation{%
  \institution{Purdue University}
  \city{Lafayette}
  \state{Indiana}
  \country{USA}
  \postcode{47901}
}
\author{Tiark Rompf}
\email{tiark@purdue.edu}
\affiliation{%
  \institution{Purdue University}
  \city{Lafayette}
  \state{Indiana}
  \country{USA}
  \postcode{47901}
}

\begin{abstract}
  \rhyme
\end{abstract}

\keywords{}

\maketitle

\section{Introduction}
\subsection{Paper Structure}
\section{Conclusion}
%\newpage
\bibliographystyle{acm}
\bibliography{paper}
\end{document}
\endinput